Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-18 ind_files/99_Appendix_D_examples.tex
Correct some refs that pointed to the wrong eqns in the fourier section

In Chapter \ref{chap_infD_processes}, we used various approximations to arrive at the linear functional Fokker-Planck equation
\begin{equation}
\label{functional_WNE_chap_4}
    \frac{\partial P}{\partial t}(\zeta,t) = \int\limits_{\mathcal{T}}\left(-\frac{\delta}{\delta \zeta(x)}\left\{\mathcal{D}_{\zeta}[\mathcal{A}^{-}](x)P(\zeta,t)\right\}+\frac{1}{2}\mathcal{A}^{+}(x|\psi)\frac{\delta^2}{\delta\zeta(x)^2}\{P(\zeta,t)\}\right)dx
\end{equation}
for describing stochastic fluctuations $\zeta$ from the deterministic solution obtained by solving \eqref{deterministic_traj}. Our goal is now to find a method to effectively detect and describe evolutionary branches (modes in trait space, corresponding to individual morphs) for this process. Following the methods used by Tim Rogers and colleagues for various special cases~\citep{rogers_demographic_2012, rogers_spontaneous_2012, rogers_modes_2015}, we will do this in a general manner by measuring the autocorrelation of the distribution of the population over trait space, a task made easier by moving to Fourier space.
\myfig{0.8}{figures/App_fourier_fig.png}{\textbf{Schematic description of Fourier analysis}. A function $\phi(x)$ (shown in red) over the trait space can be decomposed as the sum of infinitely many Fourier modes (shown in blue) $\phi_k$. In the Fourier dual space, we can look at the peaks of each of these Fourier modes: The magnitude of $\phi_k$ tells us how much it contributes to the actual function of interest $\phi$.}{fig_Fourier}
Specifically, a convenient theorem due to Weiner and Khinchin relates the autocorrelation of a probability distribution to its power spectral density via Fourier transformation. This has been extensively used in spatial ecology, and we too will make use of it here. We will thus restrict ourselves to cases in which we can express our focal function $\phi$ in terms of the Fourier basis $\{e^{ikx}\}_{k\in\mathbb{Z}}$ (Figure \ref{fig_Fourier}). For example, this can be done by restricting ourselves to cases where $\mathcal{T}$ is an interval with `periodic boundary conditions' (\emph{i.e.} we will extend all our functions from $\mathcal{T}$ to $\mathbb{R}$ in a way that they appear periodic with period given by the length of the interval $\mathcal{T}$). We may also need to restrict ourselves to a `nice' subspace of $\mathcal{M}(\mathcal{T})$, for example by intersecting with $L^2(\mathcal{T})$. In any case, we will assume all the prerequisites required for a Fourier basis expansion are satisfied. If $\mathcal{D}_{\zeta}[\mathcal{A}^{-}]$ is a translation-invariant\footnote{This is horrible nomenclature by the mathematicians. Though `invariant' is the conventional name for this concept, the intended meaning is not really invariant but `equivariant'. Formally, let $\mathcal{F}$ be a suitable function space of real valued functions. For any $c \in \mathbb{R}$, let $T_c: \mathcal{F} \to \mathcal{F}$ be the translation operator on this space, defined by $T_c[f(x)] = f(x+c)$. An operator $L: \mathcal{F} \to \mathcal{F}$ is said to be translation-invariant if it commutes with $T_c$ for every $c \in \mathbb{R}$, \emph{i.e.} $T_c[L[f]] = L[T_c[f]] \ \forall \ f \in \mathcal{F} \ \forall \ c \in \mathbb{R}$.} linear operator, then $\exp(ikx)$ acts as an eigenfunction, significantly simplifying the calculations. We therefore assume that $\mathcal{D}_{\zeta}[\mathcal{A}^{-}]$ takes the form:
\begin{equation*}
 \mathcal{D}_{\zeta}[\mathcal{A}^-](x,t) = L[\zeta(x,t)]   
\end{equation*}
for a translation-invariant linear operator $L$ that only depends on $x$ and $t$. This is not as restrictive as it initially sounds. For example, both the Laplacian operator and the convolution operator are linear and translation invariant. The presence of phenotypic clustering and polymorphisms can be analyzed by examining the power spectrum of $\Tilde{P}_{0}(\zeta,s)$ over the trait space, which is precisely what we will do.

@@

Using \eqref{fourier_representations_functions}, \eqref{fourier_representations_derivative}, and \eqref{fourier_representation_linear_operator} in \eqref{functional_WNE_zeroth_order}, we get, for the first term of the RHS:

@@

Substituting \eqref{fourier_FPE_first_term} and \eqref{fourier_FPE_second_term} into \eqref{functional_FPE}, we see that the Fokker-Planck equation in Fourier space reads:

>>>>

In Chapter \ref{chap_infD_processes}, we used various approximations to arrive at the linear functional Fokker-Planck equation
\begin{equation}
\label{functional_WNE_for_App}
    \frac{\partial P}{\partial t}(\zeta,t) = \int\limits_{\mathcal{T}}\left(-\frac{\delta}{\delta \zeta(x)}\left\{\mathcal{D}_{\zeta}[\mathcal{A}^{-}](x)P(\zeta,t)\right\}+\frac{1}{2}\mathcal{A}^{+}(x|\psi)\frac{\delta^2}{\delta\zeta(x)^2}\{P(\zeta,t)\}\right)dx
\end{equation}
for describing stochastic fluctuations $\zeta$ from the deterministic solution obtained by solving \eqref{deterministic_traj}. Our goal is now to find a method to effectively detect and describe evolutionary branches (modes in trait space, corresponding to individual morphs) for this process. Following the methods used by Tim Rogers and colleagues for various special cases~\citep{rogers_demographic_2012, rogers_spontaneous_2012, rogers_modes_2015}, we will do this in a general manner by measuring the autocorrelation of the distribution of the population over trait space, a task made easier by moving to Fourier space.
\myfig{0.8}{figures/App_fourier_fig.png}{\textbf{Schematic description of Fourier analysis}. A function $\phi(x)$ (shown in red) over the trait space can be decomposed as the sum of infinitely many Fourier modes (shown in blue) $\phi_k$. In the Fourier dual space, we can look at the peaks of each of these Fourier modes: The magnitude of $\phi_k$ tells us how much it contributes to the actual function of interest $\phi$.}{fig_Fourier}
Specifically, a convenient theorem due to Weiner and Khinchin relates the autocorrelation of a probability distribution to its power spectral density via Fourier transformation. This has been extensively used in spatial ecology, and we too will make use of it here. We will thus restrict ourselves to cases in which we can express our focal function $\phi$ in terms of the Fourier basis $\{e^{ikx}\}_{k\in\mathbb{Z}}$ (Figure \ref{fig_Fourier}). For example, this can be done by restricting ourselves to cases where $\mathcal{T}$ is an interval with `periodic boundary conditions' (\emph{i.e.} we will extend all our functions from $\mathcal{T}$ to $\mathbb{R}$ in a way that they appear periodic with period given by the length of the interval $\mathcal{T}$). We may also need to restrict ourselves to a `nice' subspace of $\mathcal{M}(\mathcal{T})$, for example by intersecting with $L^2(\mathcal{T})$. In any case, we will assume all the prerequisites required for a Fourier basis expansion are satisfied. If $\mathcal{D}_{\zeta}[\mathcal{A}^{-}]$ is a translation-invariant\footnote{This is horrible nomenclature by the mathematicians. Though `invariant' is the conventional name for this concept, the intended meaning is not really invariant but `equivariant'. Formally, let $\mathcal{F}$ be a suitable function space of real valued functions. For any $c \in \mathbb{R}$, let $T_c: \mathcal{F} \to \mathcal{F}$ be the translation operator on this space, defined by $T_c[f(x)] = f(x+c)$. An operator $L: \mathcal{F} \to \mathcal{F}$ is said to be translation-invariant if it commutes with $T_c$ for every $c \in \mathbb{R}$, \emph{i.e.} $T_c[L[f]] = L[T_c[f]] \ \forall \ f \in \mathcal{F} \ \forall \ c \in \mathbb{R}$.} linear operator, then $\exp(ikx)$ acts as an eigenfunction, significantly simplifying the calculations. We therefore assume that $\mathcal{D}_{\zeta}[\mathcal{A}^{-}]$ takes the form:
\begin{equation*}
 \mathcal{D}_{\zeta}[\mathcal{A}^-](x,t) = L[\zeta(x,t)]   
\end{equation*}
for a translation-invariant linear operator $L$ that only depends on $x$ and $t$. This is not as restrictive as it initially sounds. For example, both the Laplacian operator and the convolution operator are linear and translation invariant. The presence of phenotypic clustering and polymorphisms can be analyzed by examining the power spectrum of $\Tilde{P}_{0}(\zeta,s)$ over the trait space, which is precisely what we will do.

@@

Using \eqref{fourier_representations_functions}, \eqref{fourier_representations_derivative}, and \eqref{fourier_representation_linear_operator} in \eqref{functional_WNE_for_App}, we get, for the first term of the RHS:

@@

Substituting \eqref{fourier_FPE_first_term} and \eqref{fourier_FPE_second_term} into \eqref{functional_WNE_for_App}, we see that the Fokker-Planck equation \eqref{functional_WNE_for_App} in Fourier space reads:
